The Folly of Passwords

### 2014-12-06 paper.tex
First pass of chaper 2 done.

\section{Password Strength}

@@

@@

@@

>>>>

\section{Password Strength}
\label{sec:strength}

@@

Brute-force password guessing can also be mitigated by
users changing passwords frequently enough that attackers
have a short window within which to guess a particular
password.

As with password strength, this is only mitigating the
brute-force guessing attack vector as phishing, keylogging, etc.
can still give an attacker a password for a short while -- until
the password is next changed. As with password strength, this
is also most effective against offline brute-force attacks (e.g.
the attacker has acquired a list of hashed passwords) an
account locking policy also prevents an online brute-force
attack without the need for passwords to expire.

It is a fairly common practice in many organisations to
implement a password expiration policy whereby the
authentication systems will require a user to change their
password after some defined period of time.

\cite{adams1999users} note that password expiry increases
the burden on users having to memorise multiple passwords and that
50\% of users in their study wrote passwords down, with one
person citing a monthly password expiration policy as the cause
of having to do so.

\cite{zhang2010security} go further to demonstrate the ease of
algorithmically predicting a user's new password when the old,
expired password is known. They assert that password expiration
is weak at achieving the intent behind its implementation and
support the idea of moving away from passwords entirely.

@@

Security advice given to users usually includes the suggestion that users
have a unique password for each service they use. This can
be an unreasonable burden on users\parencite{florencio2014password}
and is thus an unrealistic expectation for an organisation or online
service.

\cite{preibusch2010password} showed how the pressure to reuse passwords
presents negative externalities from security-indifferent websites upon
security-concerned websites. This presents an organisation with
security risks stemming solely from the existence of other, less
secure websites with which its employees have accounts.

The secret nature of passwords means an organisation cannot \emph{enforce}
the use of unique passwords, so it must either make reuse impossible or
reduce the impact from reuse. Businesses can make reuse impossible
by having password strength requirements that are stronger than
services less concerned with security, with which employeess may also have
accounts. This is susceptible to the issues in section~\ref{sec:strength}
where this increases costs on users over and above than those they will
suffer from a security breach.

Clearly, password reuse is a coping strategy for users arising from the
difficulty humans have with passwords as an authentication mechanism. This
adds weight further to the case for a real shift away from passwords in
general. We will explore this case in section~\ref{sec:case} and evaluate
the alternatives to password use in chapter~\ref{chapter:alternatives}.

@@

In this chapter, we have built up the following claims:

\begin{enumerate}
  \item Strong passwords reduce guessability, but do not protect against
    many other attacks whilst adding an economically inefficient cost to users'
    time and effort.
  \item Password expiration policies are similar in only protecting against
    brute-force guessing and their effectiveness is called into question when
    users use simple mutations to choose their next password.
  \item Organisations and websites do not exist in isolation, but instead in
    an ecosystem where their users have passwords on other services and are likely
    to reuse them to reduce the burden of memorising too many.
\end{enumerate}

In summary, the use of passwords is problematic for organisations and policies
to reduce those problems place the costs on the wrong people: the users. We can
make the case here for a shift away from passwords in general or at least
for further research into user-centric designs that handle the costs of security
rationally with suitable consideration given to usability.

In chapter~\ref{chapter:alternatives}, we look at some alternatives to passwords
and evaluate them.
